---
title: "Lab2"
author: "Lydia Stephens"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading in packages and data
##packages
```{r}
library(tidyverse)
library(sf)
library(tmap)
library(stringr)
library(matrixStats)
```
##data
```{r}

counties <- sf::read_sf("./data/CBW/County_Boundaries.shp") %>% 
  sf::st_make_valid()
  
dams <- sf::read_sf("./data/CBW/Dam_or_Other_Blockage_Removed_2012_2017.shp") %>% 
  sf::st_make_valid()
  
streams <- sf::read_sf("./data/CBW/Streams_Opened_by_Dam_Removal_2012_2017.shp") %>% 
  sf::st_make_valid()

bmps <- read_csv("./data/CBW/BMPreport2016_landbmps.csv")

```

#Task 1
##1.1- Summary Statistics for cost of BMP's by state
```{r}

bmps %>% group_by(StateAbbreviation) %>% na.omit(.,) %>% summarise(Total = sum(Cost), Average = mean(Cost), StDev = sd(Cost), Median = median(Cost), Max = max(Cost), Min = min(Cost))

```
##1.2- Scatterplot Cost vs. TotalAmountCredited (Acres)
```{r}
bmps %>% dplyr::filter(Unit == "Acres") %>% ggplot(., aes(x = TotalAmountCredited, y = Cost)) +
  geom_point() +
  labs(title = "Total Amount Credited vs. Cost by State",
x = "Total Amount Credited (ln($))",
y = "Cost (ln($))") +
  scale_y_log10() +
  scale_x_log10() +
  theme_minimal()
```
##1.3-Boxplot of cover crop BMPs
```{r}
bmps %>% dplyr::filter(str_detect(BMP, "Cover Crop") & TotalAmountCredited < 5000) %>%
ggplot(., aes(x = StateAbbreviation, y = TotalAmountCredited)) +
  geom_boxplot() +
  labs(title = "Total Amount Credited for Cover Crop BMP Projects Across States",
x = "State",
y = "Total Credit (ln($))") +
  scale_y_log10() +
  theme_minimal()

```
##1.4-Dam scatterplot
```{r}
dams %>% dplyr::filter(YEAR > 0) %>% ggplot(., aes(x = YEAR, y = STATE)) +
  geom_point() +
  labs(title = "Timeline of Dam Implementation Across Virginia and Pennsylvania", x = "Year", y = "State") +
  theme_minimal()
```
##1.5 Joined data analysis
BMP in waterbodies affected by dams
```{r}
bmp.dam <- left_join(bmps, dams, by = c("StateAbbreviation" = "STATE"))


bmp.dam %>% drop_na(HUC6_NAME) %>%
ggplot(., aes(x = HUC6_NAME, fill = BMPType, na.rm = TRUE)) +
geom_bar() +
labs(title = "BMP Type in Dam-Affected Water Basins",
x = "Basin Name",
y = "Frequency",
fill = "BMP") +
theme(axis.text = element_text(angle = 60, hjust = 1))

```
#Task 2
##2.1- 5 longest rivers in 'dams' dataset
```{r}
streams %>% slice_max(LengthKM, n = 5)

```
##2.2- 3 counties with greatest total length of streams
```{r}
streams <- st_transform(streams, st_crs(counties))

streamcounty <- st_join(streams, counties, join = st_within)

streamcounty %>% group_by (NAME10) %>% summarise(StreamSum = sum(LengthKM)) %>% slice_max(StreamSum, n = 3)

```
##2.3- County map of total BMPs funded
```{r}
bmps <- bmps %>% mutate(County = str_extract(Geography, "^[^,]+"))

bmp.county <-left_join(counties, bmps, by = c("NAME10" = "County"))
bmp.county <- bmp.county %>% group_by(NAME10) %>% mutate(countytotal = sum(Cost, na.rm = T))

tm_shape(bmp.county) + tm_polygons(fill = "countytotal")
      
```
##2.4- Closest stream segment to each removed dam
```{r}
st_transform(streams, st_crs(dams))

dist <- st_nearest_feature(dams, streams)
dams <- dams %>% as.data.frame() %>% mutate(closeststream = dist) %>% left_join(dams, streams, by = c("closeststream" = "OBJECTID_1"))
view(dams)
```
##2.5- Dams per state
```{r}
dams <- sf::read_sf("./data/CBW/Dam_or_Other_Blockage_Removed_2012_2017.shp") %>% 
  sf::st_make_valid()
dams %>% group_by(STATE) %>% summarise(Removed_Dams = n())
```
