---
title: "Lab3"
author: "Lydia Stephens"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Required
##Packages
```{r}
library(spdep)
library(sf)
library(tidyverse)
library(tmap)
```
##Data
```{r}
data <-  sf::read_sf("./ACS_2020_5YR_COUNTY.gdb", layer = "ACS_2020_5YR_COUNTY")
glimpse(data)

datable <- sf::read_sf("./ACS_2020_5YR_COUNTY.gdb", layer = "X01_AGE_AND_SEX") %>%
 mutate(fixed_geoid = str_sub(GEOID, start = 8, end =-1))

datajoin <- data %>% left_join(., datable, by = c("GEOID" = "fixed_geoid"))


```
#Task 1
Create a subset of 4-7 states and save them in a shapefile
```{r}
states <- sf::read_sf("./data/statesub.shp")

```
#Task 2
Choose a variable and normalize if necessary
```{r}
states <- states %>% mutate(under5 = ((B01001e3 + B01001e27)/B01001e1), count5 = (B01001e3 + B01001e27))
```
Males and Females under 5 years old normalized by total 'sex by age' population
#Task 3
Histogram of chosen variable
```{r}
states %>% 
ggplot(., aes(x = under5, na.rm = TRUE)) +
geom_histogram() +
labs(title = "Distribution of Children Under 5 Across U.S. Northwest Counties",
x = "Proportion of Population Under 5",
y = "Frequency") 
```
#Task 4
Choropleth map of chosen variable
```{r}
tmap_mode("view")
tm_shape(states) +
  tm_polygons("under5", popup.vars = c("Number_Under_5: " = "count5", "Population: " = "B01001e1", id = "NAMELSA"), title = "Proportion of Population Under 5") +
tm_layout(title = "Distribution of Children Under 5 Across the Northwest U.S.")
```
#Task 5
Contiguous spatial weights matrix (rook or queen)
##5.1
Row-standardize the W
```{r}
state.projected <- states %>% sf::st_transform(., "ESRI:102010")

neigh <- spdep::poly2nb(state.projected, queen = TRUE)

rowstand <- nb2listw(neigh, style="W", zero.policy=TRUE)

```
##5.2
Histogram of # of neighbors
```{r}
neighbors <- attr(rowstand$weights,"comp")$d
hist(neighbors)
```
##5.3
Calculate average # of neighbors
```{r}
AU5.lag <- lag.listw(rowstand, state.projected$under5)
AU5.lag
```
##5.4
Make a Moran Plot
```{r}
moran.test(state.projected$under5, rowstand)
moran.plot(state.projected$under5, rowstand, zero.policy=TRUE, plot=TRUE)
```
#Task 6
Contiguous spatial weights matrix using IDW method (spdep)
```{r}
centroids <- st_centroid(states)
distneigh <- knn2nb(knearneigh(centroids, k=5))
id.neigh <- nb2listwdist(distneigh, centroids, style="W", type = "idw",
 alpha = 2, zero.policy = TRUE)

distneighbors <- attr(id.neigh$weights,"comp")$d
hist(distneighbors)

distAU5.lag <- lag.listw(id.neigh, state.projected$under5)
distAU5.lag

 moran.test(state.projected$under5, id.neigh)
 
 moran.plot(state.projected$under5, id.neigh, zero.policy=TRUE, plot=TRUE)
```
#Questions
##1 How is Moran's I calculated?
Moran's I is calculated by comparing neighbors and their variable values (e.g. proportion under 5). If neighbors are not similar to each other, it produces a number closer to -1, as there would be independent dispersion. On the other hand, if values cluster together, it produces a number closer to +1. If the values of the matrix were random (i.e. neither highly dispersed or clustered), the value would be 0.

##2 What is a spatially-lagged variable?
Spatially lagged variables are variables in spatial correlation models that are influenced by their neighbors. This means that polygons/points that are nearby each other are likely to have more similar values.
##3 How does your lab analysis vary across W formalization methods?
Using a k value of 5, which was similar to the average number of contiguous numbers, the Moran test results were extremely similar, with the only noticeable difference being a slightly larger range in the IDW method. Where the two methods differed was in neighbor distribution. The contiguous queen method resulted in an average of ~5 neighbors with a fairly normal distribution. The IDW method also peaked ~5 neighbors, but there was a wide spread of values. This may contribute to false results as points with high numbers of neighbors are likely going to cause clustering or dispersal based on their value.
##4 What does it mean if an observation falls in the "H-L" quadrant and why is it important to detect?

Outliers where they have a higher value than their surrounding neighbors. These outliers are important to detect because it can indicate a large disparity that needs to be addressed (e.g. natural disaster risk or hazardous waste exposure).
